---
title: "Lake Mixing Eddie Module"
output: html_notebook
---

# Eddie Module Lake Mixing  
This is an adaption of the eddie module on lake mixing and is here primarily to do the graphing of the air temperature, wind speed and Schmidts stability   
     
    
## Install libraries     
Install  libraries for use to examine data using tidyverse this only has to be done one time and a # can be put in front of the code to comment it out so it does not run in the future as now.   
```{r install libraries}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("janitor")
# install.packages("lubridate")
# install.packages("broom")
# install.packages("patchwork")
```

## Load libraries    
Load the libraries each time you run a script or program   
```{r load libraries}
library(tidyverse) # loads a lot of libraries in this one to add more functionality
library(readxl) # allows you to read in excel files
library(janitor) # cleans up column names and removes empty columns if wanted
library(lubridate) # allows easy conversion of varaibles to date format
library(broom) # cleans up output for easy presentation
library(scales) # works with the x and y scales on the plots
library(patchwork) # allows you to plot several graphs on one page
```
   
## Read file   
Note since they are all in different sheets you can change         
      sheet = "name of lake"     
to the name you are interested in to read in that lake       

Read in data on the max dephts, wind, temp and stability

```{r read data}
st_depth.df <- read_excel("../data/plotting of max st and depth.xlsx")
                         # identifies the file to read in
```

## Plot of maximum stability and depth
This is the plot of maximum stabilty a dependent variable versus the independent varaible of depth

```{r}
st_depth.df %>%
  ggplot(aes(x = max_depth, y = max_st, color = mixing_pattern)) + 
  # note that color = mixing_pattern adds a color by mixing pattern 
  # try changing this to lake as below
  geom_point(size = 4)
```


```{r}
st_depth.df %>%
  ggplot(aes(x = max_depth, y = max_st, color = lake)) + 
  # note that color = mixing_pattern adds a color by mixing pattern 
  # try changing this to lake as below
  geom_point(size = 4)
```